Towards planning of infantile cardiac surgery based on simulation

### 2012-02-24 Doc/Miccai/miccai2012.tex
Added introduction to paper.
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@331 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\section{Introduction}
State of the Art

Figure illustrating a surgery

Requirements: Model blood vessels/wall deformation, surgical low-level tasks

Plan

\section{Blood Vessel Model} % {{{

>>>>

\section{Introduction}

TODO: Include only a few sentences to CHD in general.

Surgical interventions in infants with congenitally malformed great arteries and hearts can be difficult due to their disease pattern's complex and heterogenous nature. Today, patient-specific 3-dimensional heart models can be acquired by modern techniques of non-invasive imaging and image post-processing [REFERENCE]. Nevertherless, cardiac surgeons tend to rely solely on subsets or 2-dimensional views of obtained 3-dimensional image data for examination. Possible appropriate surgical procedures may be sketched conceptually and the most promising approach is chosen just in time during the actual open-heart operation. This is the first time the surgeon gets a concrete idea of possible outcomes for different surgical procedures. However, his dicision has to be made in relatively short time and can still be based heavily on imagination capabilities and experience.

To improve planning of these kind of cardiac surgeries, optained 3-dimensional heart models could be used as input for patient-specific surgical simulation systems. Present simulation systems focus on low-level surgical procedures [REFERENCES], i.e. performing actual incisions and suturing with special input devices in real-time. Their purpose is either by means of educational training or planning access strategies to reach the heart's or blood vessel's malformations. It is more important in such \emph{reactive} simulation systems to enable simulations in real-time than to simulate tissue and blood vessels in a physically accurate manner.

To be able to predict the result of complex surgical procedures, a physically correct simulation model must be used. [REFERENCE] present a \emph{predictive} surgical simulation system for bidirectional Glenn shunts and basically similar surgeries. They use a hybrid model to simulate deformations of blood vessel trees by binding a surface mesh to a reference Cosserat rod elastically. While the mesh models surface details of blood vessels, the Cosserat rod models global deformations, i. e. stretching, bending, twisting and shearing. Low-level surgical procedures are done automatically by the software with only a minimum of interaction required which allows the surgeon to focus on the actual overall surgical procedure. Furthermore the authors show that they can attach triangular shaped patches at loose ends of dissected vessels to allow joining of vessels with different circumferences. Given the input of the surgeon, a typical simulation takes only seconds to produce a predictive result for further examination and evaluation of possibly undesired collisions or deformations of blood vessels. Although the results are promising, the author's method is limited to uniform tubular structures and operations on them, i. e. joining blood vessels end-to-end or end-to-side. Connection of blood vessels is done by using springs. This approach introduces surplus energy to the physical system and special care has to be taken during the simulation to compensate for that. Surgical procedures heavily changing surface details of blood vessels, i. e. arbitrary shaped patches to correct complicated coarctations and stenosis, are not possible with this hybrid model because the blood vessel's physical behavior is simulated by the Cosserat rod and its \emph{binding} to the surface mesh. The authors state that more sophisticated physical models including finite element methods and thin shell elements would allow this but are computationally too expensive to produce results in reasonable time.

In this paper we present an alternative, universal approach using a simulation model solely based on elastic thin shell elements. Our method is not limited to well-formed blood vessels and captures all low-level surgical procedures necessary for aortoplasty. By using a relaxation process we arbitrarily join or attach blood vessels and patches anywhere without using springs that would add surplus energy to the physical system. We prove that simulation time can be expected to be still reasonable or even decreased by taking special care of the blood vessel's overall tubular structure during generation of thin shell elements.

TODO: Structure of Paper.

\section{Blood Vessel Model}
